---
title: "WhoGov_script"
author: "Domenic Hög"
date: "2024-06-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading data set

```{r}
WhoGov_within <- read.csv("WhoGov_within_V2.0.csv")
```

# Data wrangling
## Splitting names 
```{r}
WhoGov_within <- WhoGov_within %>%
  separate(name, into = c("First Name", "Last Name"), sep = " (?=[^ ]+$)", remove = FALSE)
```

## Deleting special characters
```{r}
library(stringr)

replace_special_chars <- function(x) {
  str_replace_all(x, c(
    "\\(" = " ",   
    "\\)" = " ",   
    "&" = "and",   
    "_" = " ",     
    "," = " ",     
    "'" = ""      
  ))
}

WhoGov_within <- WhoGov_within %>%
  mutate_all(~ replace_special_chars(.))

# Forgotthe hyphens
replace_hyphens <- function(x) {
  str_replace_all(x, "-", " ")
}

WhoGov_within <- WhoGov_within %>%
  mutate_all(~ replace_hyphens(.))

# Replacing the Min. by Minister
replace_min<- function(x) {
  str_replace_all(x, "Min\\.", "Minister")
}

WhoGov_within <- WhoGov_within %>%
  mutate_all(~ replace_min(.))

```


# Preliminary descriptive statistics
## Creating data grouped by country 
```{r}
WhoGov_grouped <- group_by(WhoGov_within, country_isocode) %>%
  summarise(missing_birthyear = sum(is.na(birthyear)),
    actual_birthyear = sum(!is.na(birthyear)),
    missing_deadyear = sum(is.na(deadyear)),
    actual_deadyear = sum(!is.na(deadyear)),
    missing_name = sum(is.na(name)),
    actual_name = sum(!is.na(name)),
    missing_gender = sum(is.na(gender)),
    actual_gender = sum(!is.na(gender)),
    missing_position = sum(is.na(position)),
    actual_position = sum(!is.na(position)),
    missing_party = sum(is.na(party) | party == "unknown"),
    actual_party = sum(!is.na(party) & party != "unknown"),
    missing_portfolio_1 = sum(is.na(portfolio_1)),
    actual_portfolio_1 = sum(!is.na(portfolio_1))) %>%
  mutate(percent_birthyear = (actual_birthyear / (actual_birthyear + missing_birthyear)) * 100,
    percent_deadyear = (actual_deadyear / (actual_deadyear + missing_deadyear)) * 100,
    percent_name = (actual_name / (actual_name + missing_name)) * 100,
    percent_gender = (actual_gender / (actual_gender + missing_gender)) * 100,
    percent_position = (actual_position / (actual_position + missing_position)) * 100,
    percent_party = (actual_party / (actual_party + missing_party)) * 100,
    percent_portfolio_1 = (actual_portfolio_1 / (actual_portfolio_1 + missing_portfolio_1)) * 100
  )

print(WhoGov_grouped)

```

Especially for the birthyear and deadyear variables the data coverage is not always perfect, so that in several countries the value is below 20 percent. Party and portfolio are better coverage and mostly at least above 75% coverage. 
