---
title: "Data cleaning_NELDA"
author: "Domenic Hög"
date: "2024-06-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading NELDA data

```{r}
NELDA_broad <- read_xls("NELDA.xls")
```

## Description: steps to go in Nelda
- check if special characters used 
- changing date structure (day month year in different collumns)
- Test_merge with WhoGov

## Deleting special characters 
Deleted special characters are - / and '. Besides, I have decoded N/A in the df to NA so that they are recognised as missing values. 
```{r}
library(stringr)
library(dplyr)

NELDA_broad <- NELDA_broad %>%
  mutate(across(where(is.character), ~ na_if(., "N/A")))

NELDA_broad <- NELDA_broad %>%
  mutate(across(where(is.character), ~ str_replace_all(., "-", " ")))

NELDA_broad <- NELDA_broad %>%
  mutate(across(where(is.character), ~ str_replace_all(., "/", " ")))

NELDA_broad <- NELDA_broad %>%
  mutate(across(where(is.character), ~ str_replace_all(., "'", "")))

```


### Checking for missing values and included countries 
```{r}
prop.table(table(NELDA_broad$nelda30, useNA = "ifany"))
prop.table(table(NELDA_broad$nelda30))
table(NELDA_broad$nelda45, useNA = "ifany")
prop.table(table(NELDA_broad$nelda45, useNA = "ifany"))
prop.table(table(NELDA_broad$nelda45))
table(NELDA_broad$nelda48, useNA = "ifany")
prop.table(table(NELDA_broad$nelda48, useNA = "ifany"))
prop.table(table(NELDA_broad$nelda48))
```
### Tests in Nelda for country availabiltiy 
```{r}
length(unique(NELDA_broad$stateid))

6*31+9
```

